---
title: "4. OTU Workflow"
description: |
  Workflow for processing 16S rRNA samples for OTU analysis using mothur. Workflow uses paired end reads, beginning with raw fastq files, ending with sequence and taxonomy tables. A Microtable Object is produced to collate the data for downstream analysis. 
listing: 
    id: otu-listing
    contents: data-otu.yml
    type: table
    sort-ui: false
    filter-ui: false
    fields: 
      - filename
      - description
    field-links: 
      - filename
    field-display-names: 
      filename: File Name
      description: Description
---

{{< include /include/_setup.qmd >}}

```{r}
#| echo: false
#| eval: true
#| comment: set wd for here
remove(list = ls())
workflow_name <- "otu" # e.g., med, dada2, etc
work_here <- paste0("working_files/ssu/", workflow_name)
share_here <- paste0("share/ssu/", workflow_name)
source(here("assets/scripts", "summarize_objs.R"))
source(here("assets/scripts", "mia_metrics.R"))
```

## Workflow Input

::: {.callout-note icon=false}
## Data & Scripts

Fastq sequence files, scripts, and other assets for running this workflow can be found on the [OTU Data Portal](/workflows/portal/data-otu.qmd) page. 

The Data Portal page also contains a link to the curated output of this pipeline–-feature-sample count table, taxonomy table, sample metadata, & representative fasta sequences. Data is available as stand-alone text files, OR bundled in microtable and phyloseq objects. Archive also includes a table tracking sample read changes.
:::

#### Required Packages & Software

To run this workflow you need to have [mothur installed](https://mothur.org/wiki/installation/) and you will need the [tidyverse](https://www.tidyverse.org) package. 

## Overview

This is a standard pipeline for generating OTUs using [mothur](https://mothur.org/)[@schloss2009mothur]. We generally followed the mothur [MiSeq SOP](https://mothur.org/wiki/miseq_sop/) when building this pipeline. Since this SOP is heavily annotated we will keep our comments here to a minimum. 

:::{.columns}
:::{.column width="50%"}
{{< include include/_otu_flowchart_1.qmd >}}
:::
:::{.column width="50%"}
{{< include include/_otu_flowchart_2.qmd >}}
:::
:::

# Read Processing

{{< include include/_OTU_Part1.qmd >}}

# Dataset Prep

In this next part of the workflow our main goal is to create a *microtable object* using the R package [microeco](https://joey711.github.io/phyloseq/) [@liu2021microeco]. The microtable will be used to store the OTU by sample data as well the taxonomic, fasta, and sample data in a single object. More on that in a moment.

```{r}
#| eval: true
#| echo: false
#| message: false 
#| results: hide
#remove(list = ls())
load(here("page_build", "otu_part2.rdata"))
objects()
```

## Getting Files from Mothur

To create a microtable object we need a a sequence table, taxonomy table, and a sample data table. To generate the sequence table we need a shared file from mothur, which we can generate using the command `make.shared`. The data in a shared file represent the number of times that an OTU is observed in multiple samples. 

```{verbatim}
make.shared(list=final.opti_mcc.list, count=final.count_table, label=0.03)
```

```         
0.03

Output File Names:
final.opti_mcc.shared
```
Next we use `classify.otu` to get the OTU taxonomy table. 

```{verbatim}
classify.otu(list=final.opti_mcc.list, count=final.count_table, taxonomy=final.taxonomy, label=0.03)
```

```         
0.03

Output File Names: 
final.opti_mcc.0.03.cons.taxonomy
final.opti_mcc.0.03.cons.tax.summary
```

```{verbatim}
count.groups(shared=final.opti_mcc.shared)
```

```         
Size of smallest group: 14.

Total seqs: 34611554.

Output File Names: 
final.opti_mcc.count.summary
```

## Prep Data for `microeco`

Like any tool, the microeco package needs the data in a specific form. I formatted our data to match the mock data in the microeco tutorial, specifically [this section](https://chiliubio.github.io/microeco_tutorial/basic-class.html#prepare-the-example-data).

### A. Taxonomy Table

Here is what the taxonomy table looks like in the mock data.

```{r}
#| eval: false
#| echo: true
pandoc.table(taxonomy_table_16S[10:12, 1:3], emphasize.rownames = FALSE)
```

Our taxonomy file (below) needs a little wrangling to be properly formatted. 

```{r}
#| echo: true
#| eval: false
tmp_tax <- read_delim(here(work_here, "final.opti_mcc.0.03.cons.taxonomy"), 
                      delim = "\t")
```

```{r}
#| echo: false
#| eval: true
#tax1 <- tmp_tax
pandoc.table(tax1[11:13, 1:3], emphasize.rownames = FALSE, split.tables = Inf)
```

Some fancy string manipulation...

```{r}
#| echo: true
#| eval: false
tmp_tax <- data.frame(sapply(tmp_tax, 
                             gsub, 
                             pattern = "\\(\\d+\\)", 
                             replacement = ""))
tmp_tax <- data.frame(sapply(tmp_tax, 
                             gsub, 
                             pattern = ";$", 
                             replacement = ""))
tmp_tax <- separate_wider_delim(tmp_tax, 
                              cols = Taxonomy, 
                              delim = ";", names = c(
                                "Kingdom", "Phylum", 
                                "Class", "Order", 
                                "Family", "Genus" 
                                )
                              )
tmp_tax <- data.frame(sapply(tmp_tax, gsub, 
                           pattern = "^.*_unclassified$", 
                           replacement = ""))
tmp_tax$Size <- NULL
tmp_tax <- tibble::column_to_rownames(tmp_tax, "OTU")
```

And we get this ...

```{r}
#| echo: false
#| eval: true
#tax2 <- tmp_tax
pandoc.table(tax2[11:13, 1:3], emphasize.rownames = FALSE, split.tables = Inf)
```

```{r}
#| echo: true
#| eval: false
rank_prefixes <- c(
  Kingdom = "k", 
  Phylum  = "p", 
  Class   = "c", 
  Order   = "o", 
  Family  = "f", 
  Genus   = "g"
)

tmp_tax <- tmp_tax %>%
  mutate(across(everything(), ~replace_na(., ""))) %>% 
  mutate(across(names(rank_prefixes), 
                ~ paste0(rank_prefixes[cur_column()], "__", .))) %>%
tidy_taxonomy()
```

And then this. Excatly like the mock data. 

```{r}
#| echo: false
#| eval: true
#tax3 <- tmp_tax
pandoc.table(tax3[11:13, 1:3], emphasize.rownames = FALSE, split.tables = Inf)
```

### B. Sequence Table

Here is what the sequence table looks like in the mock data.

Here is what the sequence table looks like in the mock data.

```{r}
#| eval: true
#| echo: false
pandoc.table(otu_table_16S[1:3, 1:11], emphasize.rownames = FALSE)
```

These code block will return a properly formatted sequence table. 

```{r}
#| echo: true
#| eval: false
tmp_st <- readr::read_delim(
  here(work_here, "final.opti_mcc.shared"),  delim = "\t")
```

```{r}
#| echo: true
#| eval: false
tmp_st$numOtus <- NULL
tmp_st$label <- NULL
tmp_st <- tmp_st %>%
  tidyr::pivot_longer(cols = c(-1), names_to = "tmp") %>%
  tidyr::pivot_wider(names_from = c(1))

tmp_st <- tibble::column_to_rownames(tmp_st, "tmp")
```

```{r}
#| eval: true
#| echo: false
# st <- tmp_st
pandoc.table(st[1:5, 1:3], emphasize.rownames = FALSE)
```

```{zsh}
# only need to run this if reading in processed files
# code adds a tab to the beginning of first line
sed '1s/^/\t/' tmp_final.opti_mcc.fixed.shared > final.opti_mcc.fixed.shared
```

### C. Sample Table

```{r}
#| eval: true
#| echo: false
pandoc.table(sample_info_16S[1:3, 1:4], emphasize.rownames = FALSE, split.tables = Inf)
```

No problem. 

```{r}
#| echo: true
#| eval: false
samdf <- readRDS(here("working_files/ssu/sampledata", "sample_data.rds"))

samdf <- samdf %>% tibble::column_to_rownames("SampleID")
samdf$SampleID <- rownames(samdf)
samdf <- samdf %>% dplyr::relocate(SampleID)

samdf <- samdf %>%
  dplyr::filter(
    stringr::str_starts(SampleID, "Control", negate = TRUE)
    )
```

```{r}
#| eval: true
#| echo: false
pandoc.table(samdf[1:3, 1:4], emphasize.rownames = FALSE, split.tables = Inf)
```

## Experiment-level Objects

In the following section we create `microtable` and `phyloseq` objects. These are single, consolidated experiment-level data objects that organize various data types, such as OTU tables (taxa abundance), sample metadata, taxonomic classifications, as well as phylogenetic trees and reference sequences, into one structured format. This unified approach facilitates reproducible, interactive analysis and visualization by bringing all related information together. 

We begin by creating the `microtable` and then use the function `meco2phyloseq` from the [file2meco](https://github.com/ChiLiubio/file2meco) package to create the `phyloseq` object. This way all of the underlying data is identical across the two objects. 

::: callout-note
These objects contain an OTU table (taxa abundances), sample metadata, and taxonomy table (mapping between OTUs and higher-level taxonomic classifications).
:::

```{r}
#| echo: true
#| eval: false
sample_info <- samdf
tax_tab <- tmp_tax
otu_tab <- tmp_st
```

```{r}
#| echo: true
#| eval: false
tmp_me <- microtable$new(sample_table = sample_info, 
                         otu_table = otu_tab, 
                         tax_table = tax_tab)
tmp_me
```

```         
microtable-class object:
sample_table have 1849 rows and 13 columns
otu_table have 119453 rows and 1849 columns
tax_table have 119453 rows and 6 columns
```

### D. Add Representative Sequence

We can also add representative sequences for each OTU/ASV. For this step, we use the mothur command `get.oturep`. 

```{verbatim}
get.oturep(column=final.dist, list=final.opti_mcc.list, count=final.count_table, fasta=final.fasta)
```

The fasta file it returns needs a little T.L.C.

```
You did not provide a label, using 0.03.
0.03	119453

Output File Names: 
final.opti_mcc.0.03.rep.count_table
final.opti_mcc.0.03.rep.fasta
```

For that we use a tool called SeqKit [@shen2024seqkit2] for fasta defline manipulation.

```{zsh}
seqkit replace -p "\|.*" -r '' final.opti_mcc.0.03.rep.fasta > tmp2.fa
seqkit replace -p ".*\\t" -r '' tmp2.fa > tmp3.fa
seqkit replace -p "-" -r '$1' -s -w 0 tmp3.fa > otu_reps.fasta
rm tmp*
```

```{r}
#| echo: true
#| eval: false
rep_fasta <- Biostrings::readDNAStringSet(here(work_here, "otu_reps.fasta"))
tmp_me$rep_fasta <- rep_fasta
tmp_me$tidy_dataset()
tmp_me
```

```
microtable-class object:
sample_table have 1849 rows and 14 columns
otu_table have 119453 rows and 1849 columns
tax_table have 119453 rows and 6 columns
rep_fasta have 119453 sequence
```

```{r}
#| echo: false
#| eval: false
me_raw <- microeco::clone(tmp_me)
```

## Curate the Data Set

Pretty much the last thing to do is remove  low-count samples.

### Remove Low-Count Samples

```{r}
#| echo: true
#| eval: false
threshold <- 1000
tmp_no_low <- microeco::clone(me_raw)
tmp_no_low$otu_table <- me_raw$otu_table %>%
          dplyr::select(where(~ is.numeric(.) && sum(.) >= threshold))
tmp_no_low$tidy_dataset()
tmp_no_low
```

```         
41 taxa with 0 abundance are removed from the otu_table ...
```

```{r}
#| echo: false
#| eval: false
me_no_low <- microeco::clone(tmp_no_low)
```

```{r}
#| echo: false
#| eval: true
me_no_low
```

```{r}
#| echo: true
#| eval: false
me_final <- microeco::clone(tmp_no_low)
```

Lastly, we can use the package [`file2meco`](https://github.com/ChiLiubio/file2meco) to generate a [phyloseq](https://joey711.github.io/phyloseq/) object. 

```{r}
#| echo: true
#| eval: false
ps_final <- file2meco::meco2phyloseq(me_final)
```

## Summary

Now time to summarize the data. For this we use the R package [miaverse](https://microbiome.github.io) [@felix2024mia].

```{r}
#| echo: false
#| eval: false
identical(rownames(me_raw$sample_table), colnames(me_raw$otu_table))
identical(rownames(me_final$sample_table), colnames(me_final$otu_table))
```

```{r}
#| echo: false
#| eval: false
objs <- c("me_raw", "me_final")
pipe_summary <- summarize_objs(objs)
pipe_summary$me_dataset <- c(
  "original", "final")
print(pipe_summary)
```

```{r}
#| echo: false
#| eval: true
knitr::kable(pipe_summary)
```

```{r}
#| echo: false
#| eval: false
tmp_raw_ps <- file2meco::meco2phyloseq(me_raw)
tmp_final_ps <- file2meco::meco2phyloseq(me_final)
tmp_mia_raw <- mia_metrics(tmp_raw_ps)
tmp_mia_final <- mia_metrics(tmp_final_ps)

tmp_metrics_final <- rbind(tmp_mia_raw, tmp_mia_final)
rownames(tmp_metrics_final) <- c("Start", "End")

# suppose ds_metrics_final has rownames "Start" and "End"
ds_metrics_tbl <- tmp_metrics_final %>%
  t() %>%                         # transpose: metrics in rows
  as.data.frame() %>%
  tibble::rownames_to_column("Metric") %>%
  dplyr::rename(Start = Start, End = End)

ds_metrics_tbl <- ds_metrics_tbl %>%
  mutate(
    Start = ifelse(Start %% 1 == 0,
                   formatC(Start, format = "f", digits = 0),
                   formatC(Start, format = "f", digits = 3)),
    End   = ifelse(End %% 1 == 0,
                   formatC(End, format = "f", digits = 0),
                   formatC(End, format = "f", digits = 3))
  )

ds_metrics_tbl$Metric <- c(
  "Min. no. of reads",
  "Max. no. of reads",
  "Total no. of reads",
  "Avg. no. of reads",
  "Median no. of reads",
  "Total ASVs",
  "No. of singleton ASVs",
  "% of singleton ASVs",
  "Sparsity"
)
```

```{r}
#| echo: false
#| eval: true
#| tbl-cap: "Dataset metrics before and after curation."
knitr::kable(ds_metrics_tbl, format = "markdown")
```

We started off with `r nrow(me_raw$tax_table)` ASVs and `r nrow(me_raw$sample_table)` samples. After removing low-count samples, there were `r nrow(me_final$tax_table)` ASVs and `r nrow(me_final$sample_table)` samples remaining.

```{r}
#| echo: false
#| eval: false
# helper functions
make_tables <- function(obj, prefix) {
  rc <- obj$sample_sums() %>%
    as.data.frame() %>%
    tibble::rownames_to_column("SampleID") %>%
    dplyr::rename(!!paste0(prefix, "_rc") := 2)

  asv <- obj$otu_table %>%
    t() %>%
    as.data.frame() %>%
    { rowSums(. > 0) } %>%
    as.data.frame() %>%
    tibble::rownames_to_column("SampleID") %>%
    dplyr::rename(!!paste0(prefix, "_otu") := 2)

  list(rc = rc, asv = asv)
}
```

```{r}
#| echo: false
#| eval: false
#tmp_start <- microeco::clone(me_raw)
#tmp_final <- microeco::clone(me_final)
# --- input rc file ---
tmp_rc <- readr::read_delim(here(work_here, "mothur_pipeline_read_changes.txt")) 

# --- generate tables ---
start_tbls <- make_tables(microeco::clone(me_raw), "start")
final_tbls <- make_tables(microeco::clone(me_final), "final")

# --- summary ---
curate_summary <- tmp_rc %>%
  dplyr::full_join(start_tbls$rc,  by = "SampleID") %>%
  dplyr::left_join(start_tbls$asv, by = "SampleID") %>%
  dplyr::left_join(final_tbls$rc,  by = "SampleID") %>%
  dplyr::left_join(final_tbls$asv, by = "SampleID")

removed_samples <- curate_summary %>%
  dplyr::filter(is.na(final_rc)) %>%
  dplyr::pull(SampleID)
```

We lost a total of `r length(removed_samples)` samples after curating the dataset. This includes `r length(removed_samples[grepl("^Control_", removed_samples)])` NC samples and `r length(removed_samples[!grepl("^Control_", removed_samples)])` non-NC samples.

Here is a list of non-NC samples that were removed. 

```{r}
#| message: false
#| echo: false
#| eval: true
removed_samples[!grepl("^Control_", removed_samples)]
```

# Download Results {#download-results}

Quick access to read changes through the pipeline and repseqs detected in negative control samples.

::: {#otu-listing .column-body}
:::

<!------------------------------------------------------------------------>
<!-------------------- Use this area to save things ---------------------->
<!------------------------------------------------------------------------>

<!--------------------------------------->
<!-- These chunks are for curated data -->
<!--------------------------------------->
```{r}
#| echo: false
#| eval: false
workflow_name <- "otu"
dir.create(here(share_here, paste0(workflow_name, "_curated_data")), 
           recursive = FALSE, showWarnings = TRUE)
```

```{r}
#| echo: false
#| eval: false
tmp_path <- here(share_here, paste0(workflow_name, "_curated_data/"))

files <- list(
  otu = paste0(workflow_name, "_otu_table.txt"),
  tax = paste0(workflow_name, "_tax_table.txt"),
  sample = paste0(workflow_name, "_sample_table.txt"),
  rep = paste0(workflow_name, "_rep.fasta"),
  me = paste0("me_", workflow_name, ".rds"),
  ps = paste0("ps_", workflow_name, ".rds"),
  counts = paste0(workflow_name, "_track_read_counts.txt")
)
```

```{r}
#| echo: false
#| eval: false
#---------------OTU Table--------------#
tmp_otu <- me_final$otu_table %>% 
  tibble::rownames_to_column(paste0(toupper(workflow_name), "_ID"))
write_delim(tmp_otu, paste0(tmp_path, files$otu), delim = "\t")
#---------------TAX Table--------------#
tmp_tax <- me_final$tax_table %>% 
  tibble::rownames_to_column(paste0(toupper(workflow_name), "_ID"))
write_delim(tmp_tax, paste0(tmp_path, files$tax), delim = "\t")
#---------------SAMP Table--------------#
write_delim(me_final$sample_table, paste0(tmp_path, files$sample), delim = "\t")
#---------------REP FASTA--------------#
write.fasta(
  sequences = as.list(as.character(me_final$rep_fasta)),
  names = names(me_final$rep_fasta),
  file.out = paste0(tmp_path, files$rep)
)
#---------------        --------------#
saveRDS(me_final, paste0(tmp_path, files$me))
ps_final <- file2meco::meco2phyloseq(me_final)
saveRDS(ps_final, paste0(tmp_path, files$ps))
#---------------read count changes--------------#
write_delim(curate_summary, paste0(tmp_path, files$counts), delim = "\t")
file.copy(here("working_files/ssu/sampledata",  "all_metadata.txt"), 
          tmp_path, 
          overwrite = TRUE, recursive = FALSE, 
          copy.mode = TRUE, copy.date = FALSE)
```

```{r}
#| echo: false
#| eval: false
zip::zip(zipfile = here(share_here, paste0(workflow_name, "_curated_data.zip")), 
         files = here(share_here, paste0(workflow_name,"_curated_data")), 
         mode = "cherry-pick")
```


```{r}
#| echo: false
#| eval: false
file.copy(here(work_here,  "mothur_pipeline_read_changes.txt"), 
          here(share_here), 
          overwrite = TRUE, recursive = FALSE, 
          copy.mode = TRUE, copy.date = FALSE)
```

<!-------------------------------------------->
<!-- These chunks are for  processing data  -->
<!-------------------------------------------->

```{r}
#| echo: false
#| eval: false
dir.create(here(share_here, paste0(workflow_name, "_processing")), 
           recursive = FALSE, showWarnings = TRUE)
copy_here <- here(share_here, paste0(workflow_name, "_processing/"))
```

```{r}
#| echo: false
#| eval: false
#| comment: from sampledata (these data are common to all 16S wf)
tmp_sampdata_path <- "working_files/ssu/sampledata"
fs::dir_copy(path = here(tmp_sampdata_path,  "fastq_rename_results/"), 
             new_path = here(copy_here,  "fastq_rename_results/"), 
             overwrite = TRUE)

fs::dir_copy(path = here(tmp_sampdata_path,  "fastq_rename_lookup/"), 
             new_path = here(copy_here,  "fastq_rename_lookup/"), 
             overwrite = TRUE)
file.copy(here(tmp_sampdata_path,  "rename.sh"), 
          here(copy_here), 
          overwrite = TRUE, recursive = FALSE, 
          copy.mode = TRUE, copy.date = FALSE)
```


```{r}
#| echo: false
#| eval: false
fs::dir_copy(path = here(work_here,  paste0(workflow_name, "_hydra_scripts/")), 
             new_path = here(copy_here,  paste0(workflow_name, "_hydra_scripts/")), 
             overwrite = TRUE)

file.copy(here(work_here,  paste0(workflow_name, "_batchfile_processing/")), 
          copy_here, 
          overwrite = TRUE, recursive = FALSE, 
          copy.mode = TRUE, copy.date = FALSE)

file.copy(here(share_here,  "mothur_pipeline_read_changes.txt"), 
          here(copy_here), 
          overwrite = TRUE, recursive = FALSE, 
          copy.mode = TRUE, copy.date = FALSE)

file.copy(here(tmp_sampdata_path,  "all_metadata.txt"), 
          here(copy_here), 
          overwrite = TRUE, recursive = FALSE, 
          copy.mode = TRUE, copy.date = FALSE)
```

```{r}
#| echo: false
#| eval: false
zip::zip(zipfile = here(share_here, paste0(workflow_name, "_processing.zip")), 
         files = here(share_here, paste0(workflow_name,"_processing")), 
         mode = "cherry-pick")
dir_delete(here(share_here, paste0(workflow_name,"_processing")))
```

```{r}
#| echo: false
#| eval: true
#| message: false
#| include: false
#| comment: save R script
workflow_name <- "otu"
options(knitr.duplicate.label = "allow")
# Define sources and numbered outputs
sources <- c(
  "include/_OTU_Part1.qmd",
  "index.qmd"
)
# Generate corresponding temporary output files
tmp_outputs <- here("share/ssu", paste0(workflow_name, "_workflow", seq_along(sources), ".R"))
# Final combined output
final_output <- here(share_here, paste0(workflow_name, "_workflow.R"))
# Step 1: Extract code from qmd to temporary R scripts
purrr::walk2(sources, tmp_outputs, ~
  knitr::purl(.x, output = .y, documentation = 0)
)
# Step 2: Concatenate all temporary R scripts into one, with separators
sink(final_output)
purrr::walk2(tmp_outputs, seq_along(tmp_outputs), ~{
  cat(readLines(.x), sep = "\n")
  cat("\n\n",
      paste0("# -------------------- END OF SCRIPT ", .y, " --------------------\n\n"),
      sep = "")
})
sink()
# Step 3: Clean up temporary files
file.remove(tmp_outputs)
```

```{r}
#| echo: false
#| eval: false
objects()
gdata::keep(me_final, me_no_low, me_raw, curate_summary, 
            ds_metrics_tbl, pipe_summary, removed_samples, 
            samdf, st, tax1, tax2, tax3,
            sure = TRUE)
save.image(here("page_build", "otu_part2.rdata"))
```

#### References {.appendix}

::: {#refs}
:::

{{< include /include/_footer.qmd >}}

